---
title: "TIER-Plus — Scenario Comparison Report"
output:
  html_document:
    theme: null
    self_contained: true
    mathjax: null
params:
  country_name: ""
  generated_at: ""
  baseline_vals: !r list()
  scenario1_vals: !r list()
  scenario2_vals: !r list()
  outcomes_base: !r list()
  outcomes_s1: !r list()
  outcomes_s2: !r list()
  diff_s1: !r list()
  diff_s2: !r list()
  populations: !r list()
  interventions: !r list()
---

```{css, echo=FALSE}
@page {
  size: A4;
  margin: 18mm 16mm;
}

body {
  font-family: "Calibri", "Helvetica", "Arial", sans-serif;
  font-size: 10.5pt;
  color: #1a1a1a;
  line-height: 1.4;
  margin: 0;
}

h1 {
  font-size: 18pt;
  color: #1F3864;
  margin: 0 0 4px 0;
  padding-bottom: 4px;
  border-bottom: 2px solid #1F3864;
}

h2 {
  font-size: 13pt;
  color: #1F3864;
  margin: 18px 0 6px 0;
  padding-bottom: 2px;
  border-bottom: 1px solid #BFBFBF;
  page-break-after: avoid;
}

h3 {
  font-size: 11pt;
  color: #1F3864;
  margin: 12px 0 4px 0;
  page-break-after: avoid;
}

.header-meta {
  font-size: 9.5pt;
  color: #595959;
  margin-bottom: 12px;
}

.caveat {
  background-color: #fffbeb;
  border: 1px solid #f59e0b;
  border-radius: 4px;
  padding: 8px 12px;
  margin: 10px 0 14px 0;
  font-size: 9.5pt;
  color: #78350f;
}

.caveat strong { color: #92400e; }

table {
  width: 100%;
  border-collapse: collapse;
  margin: 6px 0 10px 0;
  font-size: 10pt;
  page-break-inside: avoid;
}

th, td {
  border: 1px solid #d0d0d0;
  padding: 5px 8px;
  text-align: right;
}

th:first-child, td:first-child {
  text-align: left;
}

thead th {
  background-color: #1F3864;
  color: white;
  font-weight: bold;
}

tbody tr:nth-child(even) { background-color: #f7f9fc; }

/* knitr::kable tables share these styles via the report-table class. */
table.report-table {
  width: 100%;
  border-collapse: collapse;
  margin: 6px 0 10px 0;
  font-size: 10pt;
}
table.report-table th,
table.report-table td {
  border: 1px solid #d0d0d0;
  padding: 5px 8px;
}
table.report-table thead th {
  background-color: #1F3864;
  color: white;
  font-weight: bold;
  text-align: left;
}
table.report-table thead th:not(:first-child) { text-align: right; }
table.report-table tbody td:first-child { text-align: left; }
table.report-table tbody td { text-align: right; }
table.report-table tbody tr:nth-child(even) { background-color: #f7f9fc; }
table.report-table tbody tr:last-child { background-color: #f0f4fa; }

.section-pair {
  display: flex;
  gap: 12px;
  page-break-inside: avoid;
}

.section-pair > div { flex: 1; }

.kv-row {
  display: flex;
  justify-content: space-between;
  border-bottom: 1px solid #e5e5e5;
  padding: 3px 0;
}

.kv-row .label { color: #333; }
.kv-row .value { font-weight: bold; }

.delta-positive { color: #b91c1c; }   /* worse than baseline */
.delta-negative { color: #15803d; }   /* better than baseline */
.delta-neutral  { color: #6b7280; }

.plot-container {
  page-break-inside: avoid;
  margin: 8px 0 14px 0;
}

.footer-note {
  font-size: 8.5pt;
  color: #6b7280;
  font-style: italic;
  border-top: 1px solid #d0d0d0;
  padding-top: 6px;
  margin-top: 18px;
}

/* control where we break pages */
.page-break { page-break-after: always; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      dev = "png", dpi = 150)

# Pull params into shorter names
p           <- params
base_vals   <- p$baseline_vals
s1_vals     <- p$scenario1_vals
s2_vals     <- p$scenario2_vals
out_base    <- p$outcomes_base
out_s1      <- p$outcomes_s1
out_s2      <- p$outcomes_s2
diff_s1     <- p$diff_s1
diff_s2     <- p$diff_s2
pops        <- p$populations
ig          <- p$interventions

# Formatters
fmt_n     <- function(x) format(round(as.numeric(x)), big.mark = ",", scientific = FALSE)
fmt_money <- function(x) paste0("$", fmt_n(x))
fmt_pct   <- function(x) paste0(format(round(as.numeric(x), 1), nsmall = 1), "%")

# Sign-aware delta string. good_when = "down" means a negative delta is good
# (infections, deaths, cost). "up" means a positive delta is good (ART
# initiations, suppressed).
fmt_delta <- function(value, good_when = "down", money = FALSE) {
  if (is.null(value) || is.na(value)) return("")
  formatter <- if (money) fmt_money else fmt_n
  sign_char <- if (value > 0) "+" else if (value < 0) "-" else ""
  abs_str   <- formatter(abs(value))
  cls <- if (value == 0) "delta-neutral"
         else if ((value > 0 && good_when == "up") ||
                  (value < 0 && good_when == "down")) "delta-negative"
         else "delta-positive"
  sprintf('<span class="%s">(%s%s)</span>', cls, sign_char, abs_str)
}

# 95-95-95 calculator (matches the on-screen logic)
calc_95s <- function(o) {
  d1 <- if (!is.null(o$end_plhiv)    && o$end_plhiv    > 0) o$end_diagnosed  / o$end_plhiv    * 100 else 0
  d2 <- if (!is.null(o$end_diagnosed)&& o$end_diagnosed> 0) o$end_on_art     / o$end_diagnosed * 100 else 0
  d3 <- if (!is.null(o$end_on_art)   && o$end_on_art   > 0) o$end_suppressed / o$end_on_art    * 100 else 0
  list(first = min(d1, 100), second = min(d2, 100), third = min(d3, 100))
}

g_base <- calc_95s(out_base)
g_s1   <- calc_95s(out_s1)
g_s2   <- calc_95s(out_s2)
```

<div class="header-meta">
**Regional profile:** `r p$country_name` &nbsp;|&nbsp;
**Generated:** `r p$generated_at`
</div>

<div class="caveat">
<strong>⚠ Interpretation note.</strong> This tool is <em>indicative</em>
of the direction and rough magnitude of impact of different choices. Outputs are not 
100% accurate and only treatment and included intervention costs are incorporated. 
Specific numbers used for budgeting, target-setting, or operational planning should 
come from country-led processes, supported (not replaced) by this tool. Percentages (95-95-95 targets) can also be misleading when patient numbers change — always check the absolute population numbers in the cascade chart and epidemiological outcomes before drawing conclusions.
</div>

## Progress toward 95-95-95 (end of year)

```{r tbl-95s, results='asis'}
goals_tbl <- data.frame(
  Indicator   = c("1st 95 — % of PLHIV diagnosed",
                  "2nd 95 — % of diagnosed on ART",
                  "3rd 95 — % on ART suppressed"),
  Baseline    = c(fmt_pct(g_base$first), fmt_pct(g_base$second), fmt_pct(g_base$third)),
  `Scenario 1` = c(fmt_pct(g_s1$first),  fmt_pct(g_s1$second),  fmt_pct(g_s1$third)),
  `Scenario 2` = c(fmt_pct(g_s2$first),  fmt_pct(g_s2$second),  fmt_pct(g_s2$third)),
  check.names = FALSE
)
knitr::kable(goals_tbl, format = "html", escape = FALSE,
             align = c("l", "r", "r", "r"),
             table.attr = 'class="report-table"')
```

## Key epidemiological outcomes (end of year)

```{r tbl-epi, results='asis'}
# Cell builder: value plus delta span when a delta value is supplied.
cell <- function(val, delta = NULL, good_when = "down", money = FALSE) {
  v <- if (money) fmt_money(val) else fmt_n(val)
  if (is.null(delta)) return(v)
  paste0(v, " ", fmt_delta(delta, good_when, money))
}

epi_tbl <- data.frame(
  Outcome = c("New adult acquisitions", "New infant acquisitions",
              "HIV-related deaths", "ART initiations",
              "End-of-year on ART (suppressed)"),
  Baseline = c(fmt_n(out_base$end_new_infections),
               fmt_n(out_base$end_infant_infections),
               fmt_n(out_base$end_deaths),
               fmt_n(out_base$art_initiations),
               fmt_n(out_base$end_suppressed)),
  `Scenario 1` = c(
    cell(out_s1$end_new_infections,    diff_s1$diff_new_infections,    "down"),
    cell(out_s1$end_infant_infections, diff_s1$diff_infant_infections, "down"),
    cell(out_s1$end_deaths,            diff_s1$diff_deaths,            "down"),
    cell(out_s1$art_initiations,       diff_s1$diff_art_initiations,   "up"),
    cell(out_s1$end_suppressed,        diff_s1$diff_suppressed,        "up")
  ),
  `Scenario 2` = c(
    cell(out_s2$end_new_infections,    diff_s2$diff_new_infections,    "down"),
    cell(out_s2$end_infant_infections, diff_s2$diff_infant_infections, "down"),
    cell(out_s2$end_deaths,            diff_s2$diff_deaths,            "down"),
    cell(out_s2$art_initiations,       diff_s2$diff_art_initiations,   "up"),
    cell(out_s2$end_suppressed,        diff_s2$diff_suppressed,        "up")
  ),
  check.names = FALSE
)
knitr::kable(epi_tbl, format = "html", escape = FALSE,
             align = c("l", "r", "r", "r"),
             table.attr = 'class="report-table"')
```

## Cost analysis

```{r tbl-cost, results='asis'}
cost_tbl <- data.frame(
  `Cost category` = c("Intervention costs", "ART provision", "Total programme cost"),
  Baseline = c(fmt_money(out_base$total_intervention_cost),
               fmt_money(out_base$art_provision_cost),
               paste0("<strong>", fmt_money(out_base$total_cost), "</strong>")),
  `Scenario 1` = c(
    cell(out_s1$total_intervention_cost, diff_s1$diff_intervention_cost, "down", money = TRUE),
    cell(out_s1$art_provision_cost,      diff_s1$diff_art_provision_cost, "down", money = TRUE),
    paste0("<strong>", fmt_money(out_s1$total_cost), "</strong> ",
           fmt_delta(diff_s1$diff_total_cost, "down", money = TRUE))
  ),
  `Scenario 2` = c(
    cell(out_s2$total_intervention_cost, diff_s2$diff_intervention_cost, "down", money = TRUE),
    cell(out_s2$art_provision_cost,      diff_s2$diff_art_provision_cost, "down", money = TRUE),
    paste0("<strong>", fmt_money(out_s2$total_cost), "</strong> ",
           fmt_delta(diff_s2$diff_total_cost, "down", money = TRUE))
  ),
  check.names = FALSE
)
knitr::kable(cost_tbl, format = "html", escape = FALSE,
             align = c("l", "r", "r", "r"),
             table.attr = 'class="report-table"')
```

<div class="page-break"></div>

## HIV care cascade — end of year

```{r cascade-plot, fig.width=8, fig.height=4.2, out.width="100%"}
library(ggplot2)
library(scales)

cascade_df <- data.frame(
  Stage    = rep(c("PLHIV", "Diagnosed", "On ART", "Suppressed"), times = 3),
  Value    = c(out_base$end_plhiv, out_base$end_diagnosed, out_base$end_on_art, out_base$end_suppressed,
               out_s1$end_plhiv,   out_s1$end_diagnosed,   out_s1$end_on_art,   out_s1$end_suppressed,
               out_s2$end_plhiv,   out_s2$end_diagnosed,   out_s2$end_on_art,   out_s2$end_suppressed),
  Scenario = factor(rep(c("Baseline", "Scenario 1", "Scenario 2"), each = 4),
                    levels = c("Baseline", "Scenario 1", "Scenario 2"))
)
cascade_df$Stage <- factor(cascade_df$Stage,
                           levels = c("PLHIV", "Diagnosed", "On ART", "Suppressed"))

ggplot(cascade_df,
       aes(x = Stage, y = Value, colour = Scenario, group = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.8) +
  scale_y_continuous(labels = comma) +
  scale_colour_manual(values = c("Baseline"  = "#6b7280",
                                 "Scenario 1" = "#3b82f6",
                                 "Scenario 2" = "#ef4444")) +
  labs(x = NULL, y = "Number of people") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        panel.grid.minor = element_blank())
```

## Impact summary — change from baseline

```{r impact-summary, fig.width=8, fig.height=3.6, out.width="100%"}
mk_impact_df <- function(diff_obj, scen_label) {
  data.frame(
    Outcome = c("Acquisitions", "Deaths", "ART\ninitiations"),
    Value   = c(diff_obj$diff_new_infections,
                diff_obj$diff_deaths,
                diff_obj$diff_art_initiations),
    is_good = c(diff_obj$diff_new_infections < 0,
                diff_obj$diff_deaths        < 0,
                diff_obj$diff_art_initiations >= 0),
    Scenario = scen_label
  )
}

impact_df <- rbind(mk_impact_df(diff_s1, "Scenario 1"),
                   mk_impact_df(diff_s2, "Scenario 2"))

ggplot(impact_df, aes(x = Outcome, y = Value, fill = is_good)) +
  geom_col(width = 0.65) +
  geom_hline(yintercept = 0, colour = "#374151", linewidth = 0.4) +
  facet_wrap(~ Scenario, nrow = 1) +
  scale_fill_manual(values = c("TRUE" = "#10b981", "FALSE" = "#ef4444"),
                    guide = "none") +
  scale_y_continuous(labels = comma) +
  labs(x = NULL, y = "People (relative to baseline)") +
  theme_minimal(base_size = 10) +
  theme(strip.text = element_text(face = "bold"),
        panel.grid.minor = element_blank())
```

## Intervention volumes by scenario

```{r interv-table, results='asis'}
# Build a long table of every intervention with its baseline/scenario volumes.
rows <- list()
for (gkey in names(ig)) {
  grp <- ig[[gkey]]
  # group header row
  rows[[length(rows) + 1]] <- sprintf(
    '<tr style="background-color:#eef2f7;"><td colspan="4"><strong>%s</strong></td></tr>',
    htmltools::htmlEscape(grp$name)
  )
  for (ikey in names(grp$interventions)) {
    int <- grp$interventions[[ikey]]
    bv  <- base_vals[[ikey]]; if (is.null(bv)) bv <- 0
    s1v <- s1_vals[[ikey]];   if (is.null(s1v)) s1v <- 0
    s2v <- s2_vals[[ikey]];   if (is.null(s2v)) s2v <- 0
    is_pct <- isTRUE(int$type == "coverage")
    fmt_v <- function(x) if (is_pct) sprintf("%.1f%%", x) else fmt_n(x)
    rows[[length(rows) + 1]] <- sprintf(
      "<tr><td>%s</td><td>%s</td><td>%s</td><td>%s</td></tr>",
      htmltools::htmlEscape(int$name),
      fmt_v(bv), fmt_v(s1v), fmt_v(s2v)
    )
  }
}
cat('<table>\n<thead>',
    '<tr><th>Intervention</th><th>Baseline</th><th>Scenario 1</th><th>Scenario 2</th></tr>',
    '</thead>\n<tbody>\n',
    paste(rows, collapse = "\n"),
    '\n</tbody>\n</table>',
    sep = "")
```

<div class="footer-note">
Use the tool to compare directions and trade-offs. Use country processes and
validated models for the exact numbers in your plan. Report generated by the
TIER-Plus Tool.
</div>
