---
title: "TIER-Plus — Scenario Comparison Report"
output:
  html_document:
    theme: null
    self_contained: true
    mathjax: null
params:
  country_name: ""
  generated_at: ""
  baseline_vals: !r list()
  scenario1_vals: !r list()
  scenario2_vals: !r list()
  outcomes_base: !r list()
  outcomes_s1: !r list()
  outcomes_s2: !r list()
  diff_s1: !r list()
  diff_s2: !r list()
  populations: !r list()
  interventions: !r list()
  prep_delivery: !r list()
  param_overrides: !r data.frame()
  show_art_initiations: false
---

```{css, echo=FALSE}
@page {
  size: A4;
  margin: 18mm 16mm;
}

body {
  font-family: "Calibri", "Helvetica", "Arial", sans-serif;
  font-size: 10.5pt;
  color: #1a1a1a;
  line-height: 1.4;
  margin: 0;
}

h1 {
  font-size: 18pt;
  color: #1F3864;
  margin: 0 0 4px 0;
  padding-bottom: 4px;
  border-bottom: 2px solid #1F3864;
}

h2 {
  font-size: 13pt;
  color: #1F3864;
  margin: 18px 0 6px 0;
  padding-bottom: 2px;
  border-bottom: 1px solid #BFBFBF;
  page-break-after: avoid;
}

h3 {
  font-size: 11pt;
  color: #1F3864;
  margin: 12px 0 4px 0;
  page-break-after: avoid;
}

.header-meta {
  font-size: 9.5pt;
  color: #595959;
  margin-bottom: 12px;
}

.caveat {
  background-color: #fffbeb;
  border: 1px solid #f59e0b;
  border-radius: 4px;
  padding: 8px 12px;
  margin: 10px 0 14px 0;
  font-size: 9.5pt;
  color: #78350f;
}

.caveat strong { color: #92400e; }

table {
  width: 100%;
  border-collapse: collapse;
  margin: 6px 0 10px 0;
  font-size: 10pt;
  page-break-inside: avoid;
}

th, td {
  border: 1px solid #d0d0d0;
  padding: 5px 8px;
  text-align: right;
}

th:first-child, td:first-child {
  text-align: left;
}

thead th {
  background-color: #1F3864;
  color: white;
  font-weight: bold;
}

tbody tr:nth-child(even) { background-color: #f7f9fc; }

/* knitr::kable tables share these styles via the report-table class. */
table.report-table {
  width: 100%;
  border-collapse: collapse;
  margin: 6px 0 10px 0;
  font-size: 10pt;
}
table.report-table th,
table.report-table td {
  border: 1px solid #d0d0d0;
  padding: 5px 8px;
}
table.report-table thead th {
  background-color: #1F3864;
  color: white;
  font-weight: bold;
  text-align: left;
}
table.report-table thead th:not(:first-child) { text-align: right; }
table.report-table tbody td:first-child { text-align: left; }
table.report-table tbody td { text-align: right; }
table.report-table tbody tr:nth-child(even) { background-color: #f7f9fc; }
table.report-table tbody tr:last-child { background-color: #f0f4fa; }

.section-pair {
  display: flex;
  gap: 12px;
  page-break-inside: avoid;
}

.section-pair > div { flex: 1; }

.kv-row {
  display: flex;
  justify-content: space-between;
  border-bottom: 1px solid #e5e5e5;
  padding: 3px 0;
}

.kv-row .label { color: #333; }
.kv-row .value { font-weight: bold; }

.delta-positive { color: #b91c1c; }   /* worse than baseline */
.delta-negative { color: #15803d; }   /* better than baseline */
.delta-neutral  { color: #6b7280; }

.plot-container {
  page-break-inside: avoid;
  margin: 8px 0 14px 0;
}

.footer-note {
  font-size: 8.5pt;
  color: #6b7280;
  font-style: italic;
  border-top: 1px solid #d0d0d0;
  padding-top: 6px;
  margin-top: 18px;
}

/* control where we break pages */
.page-break { page-break-after: always; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      dev = "png", dpi = 150)

# Pull params into shorter names
p           <- params
base_vals   <- p$baseline_vals
s1_vals     <- p$scenario1_vals
s2_vals     <- p$scenario2_vals
out_base    <- p$outcomes_base
out_s1      <- p$outcomes_s1
out_s2      <- p$outcomes_s2
diff_s1     <- p$diff_s1
diff_s2     <- p$diff_s2
pops        <- p$populations
ig          <- p$interventions
pd          <- p$prep_delivery
po          <- p$param_overrides

# PrEP entry mode is a single GLOBAL toggle in the app (input$prep_entry_mode),
# so baseline and both scenarios always share it -- reading one is enough.
# Defaults to by-group if prep_delivery is absent (standalone knit).
prep_mode   <- if (!is.null(pd$baseline$mode)) pd$baseline$mode else "group"
prep_total  <- identical(prep_mode, "total")

# The eight per-group PrEP keys. In total-entry mode these are DERIVED by
# split_prep_total() rather than entered, so they are collapsed to two rows in
# the intervention table (see that chunk).
PREP_GROUP_KEYS <- c("prep_oral_fsw", "prep_oral_msm", "prep_oral_agyw",
                     "prep_oral_general", "prep_lenacapavir_fsw",
                     "prep_lenacapavir_msm", "prep_lenacapavir_agyw",
                     "prep_lenacapavir_general")
PREP_ORAL_KEYS <- grep("^prep_oral_", PREP_GROUP_KEYS, value = TRUE)
PREP_LENA_KEYS <- grep("^prep_lenacapavir_", PREP_GROUP_KEYS, value = TRUE)

# ART initiations is optional -- set by SHOW_ART_INITIATIONS in the app file
# and passed through report_params. isTRUE() guards against the param being
# absent when the template is knitted standalone.
show_art    <- isTRUE(p$show_art_initiations)

# Formatters
fmt_n     <- function(x) format(round(as.numeric(x)), big.mark = ",", scientific = FALSE)
fmt_money <- function(x) paste0("$", fmt_n(x))
fmt_pct   <- function(x) paste0(format(round(as.numeric(x), 1), nsmall = 1), "%")

# Sign-aware delta string. good_when = "down" means a negative delta is good
# (infections, deaths, cost). "up" means a positive delta is good (ART
# initiations, suppressed).
fmt_delta <- function(value, good_when = "down", money = FALSE) {
  if (is.null(value) || is.na(value)) return("")
  formatter <- if (money) fmt_money else fmt_n
  sign_char <- if (value > 0) "+" else if (value < 0) "-" else ""
  abs_str   <- formatter(abs(value))
  cls <- if (value == 0) "delta-neutral"
         else if ((value > 0 && good_when == "up") ||
                  (value < 0 && good_when == "down")) "delta-negative"
         else "delta-positive"
  sprintf('<span class="%s">(%s%s)</span>', cls, sign_char, abs_str)
}

# 95-95-95 calculator (matches the on-screen logic)
calc_95s <- function(o) {
  d1 <- if (!is.null(o$end_plhiv)    && o$end_plhiv    > 0) o$end_diagnosed  / o$end_plhiv    * 100 else 0
  d2 <- if (!is.null(o$end_diagnosed)&& o$end_diagnosed> 0) o$end_on_art     / o$end_diagnosed * 100 else 0
  d3 <- if (!is.null(o$end_on_art)   && o$end_on_art   > 0) o$end_suppressed / o$end_on_art    * 100 else 0
  list(first = min(d1, 100), second = min(d2, 100), third = min(d3, 100))
}

g_base <- calc_95s(out_base)
g_s1   <- calc_95s(out_s1)
g_s2   <- calc_95s(out_s2)
```

<div class="header-meta">
**Regional profile:** `r p$country_name` &nbsp;|&nbsp;
**Generated:** `r p$generated_at`
</div>

<div class="caveat">
<strong>⚠ Interpretation note.</strong> This tool is <em>indicative</em>
of the direction and rough magnitude of impact of different choices. Outputs may not be accurate to exact numbers and changes to the percentage 95-95-95 targets should be interpreted in conjunction with the absolute change in the number of people in each group. Costs are limited to treatment and included intervention costs only. As a result, many indirect and other programme costs are excluded. Specific numbers used for budgeting, target-setting, or operational planning should come from country-led processes, supported (but not replaced) by this tool.
</div>

## Progress toward 95-95-95 (end of year)

```{r tbl-95s, results='asis'}
goals_tbl <- data.frame(
  Indicator   = c("1st 95 — % of PLHIV diagnosed",
                  "2nd 95 — % of diagnosed on ART",
                  "3rd 95 — % on ART suppressed"),
  Baseline    = c(fmt_pct(g_base$first), fmt_pct(g_base$second), fmt_pct(g_base$third)),
  `Scenario 1` = c(fmt_pct(g_s1$first),  fmt_pct(g_s1$second),  fmt_pct(g_s1$third)),
  `Scenario 2` = c(fmt_pct(g_s2$first),  fmt_pct(g_s2$second),  fmt_pct(g_s2$third)),
  check.names = FALSE
)
knitr::kable(goals_tbl, format = "html", escape = FALSE,
             align = c("l", "r", "r", "r"),
             table.attr = 'class="report-table"')
```

## Key epidemiological outcomes (end of year)

```{r tbl-epi, results='asis'}
# Cell builder: value plus delta span when a delta value is supplied.
# Shared with the cost table below (this chunk runs first).
cell <- function(val, delta = NULL, good_when = "down", money = FALSE) {
  v <- if (money) fmt_money(val) else fmt_n(val)
  if (is.null(delta)) return(v)
  paste0(v, " ", fmt_delta(delta, good_when, money))
}

# Absolute end-of-year values (with change from baseline), mirroring the
# on-screen "Key Epidemiological Outcomes" cards. ART initiations and
# suppression change are shown in the Health outcomes table below.
epi_tbl <- data.frame(
  Outcome = c("New adult acquisitions", "New infant acquisitions",
              "HIV-related deaths"),
  Baseline = c(fmt_n(out_base$end_new_infections),
               fmt_n(out_base$end_infant_infections),
               fmt_n(out_base$end_deaths)),
  `Scenario 1` = c(
    cell(out_s1$end_new_infections,    diff_s1$diff_new_infections,    "down"),
    cell(out_s1$end_infant_infections, diff_s1$diff_infant_infections, "down"),
    cell(out_s1$end_deaths,            diff_s1$diff_deaths,            "down")
  ),
  `Scenario 2` = c(
    cell(out_s2$end_new_infections,    diff_s2$diff_new_infections,    "down"),
    cell(out_s2$end_infant_infections, diff_s2$diff_infant_infections, "down"),
    cell(out_s2$end_deaths,            diff_s2$diff_deaths,            "down")
  ),
  check.names = FALSE
)
knitr::kable(epi_tbl, format = "html", escape = FALSE,
             align = c("l", "r", "r", "r"),
             table.attr = 'class="report-table"')
```

## Health outcomes (relative to baseline)

```{r tbl-health, results='asis'}
# Signed change vs baseline with % change, coloured like the on-screen
# "Health Outcomes" cards (green = better, red = worse). good_when = "up"
# means a positive change is good (ART initiations, suppression); "down"
# means a negative change is good (acquisitions, deaths). Baseline is the
# reference, so only the two scenarios are shown.
fmt_change <- function(delta, pct, good_when) {
  if (is.null(delta) || is.na(delta)) return("&mdash;")
  cls <- if (delta == 0) "delta-neutral"
         else if ((delta > 0 && good_when == "up") ||
                  (delta < 0 && good_when == "down")) "delta-negative"
         else "delta-positive"
  sign_char <- if (delta > 0) "+" else if (delta < 0) "-" else ""
  pct_str <- if (is.null(pct) || is.na(pct)) "" else sprintf(" (%+.1f%%)", pct)
  sprintf('<span class="%s">%s%s%s</span>', cls, sign_char, fmt_n(abs(delta)), pct_str)
}

# c() drops NULL, so `if (show_art)` with no `else` removes the row from all
# three columns at once and keeps the vectors the same length.
health_tbl <- data.frame(
  `Change vs baseline` = c(if (show_art) "ART initiations",
                           "Adult acquisitions",
                           "HIV-related deaths", "On ART & suppressed"),
  `Scenario 1` = c(
    if (show_art) fmt_change(diff_s1$diff_art_initiations, diff_s1$pct_art_initiations, "up"),
    fmt_change(diff_s1$diff_new_infections,  diff_s1$pct_new_infections,  "down"),
    fmt_change(diff_s1$diff_deaths,          diff_s1$pct_deaths,          "down"),
    fmt_change(diff_s1$diff_suppressed,      diff_s1$pct_suppressed,      "up")
  ),
  `Scenario 2` = c(
    if (show_art) fmt_change(diff_s2$diff_art_initiations, diff_s2$pct_art_initiations, "up"),
    fmt_change(diff_s2$diff_new_infections,  diff_s2$pct_new_infections,  "down"),
    fmt_change(diff_s2$diff_deaths,          diff_s2$pct_deaths,          "down"),
    fmt_change(diff_s2$diff_suppressed,      diff_s2$pct_suppressed,      "up")
  ),
  check.names = FALSE
)
knitr::kable(health_tbl, format = "html", escape = FALSE,
             align = c("l", "r", "r"),
             table.attr = 'class="report-table"')
```

<div class="page-break"></div>

## HIV care cascade — end of year

```{r cascade-plot, fig.width=8, fig.height=4.2, out.width="100%"}
library(ggplot2)
library(scales)

cascade_df <- data.frame(
  Stage    = rep(c("PLHIV", "Diagnosed", "On ART", "Suppressed"), times = 3),
  Value    = c(out_base$end_plhiv, out_base$end_diagnosed, out_base$end_on_art, out_base$end_suppressed,
               out_s1$end_plhiv,   out_s1$end_diagnosed,   out_s1$end_on_art,   out_s1$end_suppressed,
               out_s2$end_plhiv,   out_s2$end_diagnosed,   out_s2$end_on_art,   out_s2$end_suppressed),
  Scenario = factor(rep(c("Baseline", "Scenario 1", "Scenario 2"), each = 4),
                    levels = c("Baseline", "Scenario 1", "Scenario 2"))
)
cascade_df$Stage <- factor(cascade_df$Stage,
                           levels = c("PLHIV", "Diagnosed", "On ART", "Suppressed"))

ggplot(cascade_df,
       aes(x = Stage, y = Value, colour = Scenario, group = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.8) +
  scale_y_continuous(labels = comma) +
  scale_colour_manual(values = c("Baseline"  = "#6b7280",
                                 "Scenario 1" = "#3b82f6",
                                 "Scenario 2" = "#ef4444")) +
  labs(x = NULL, y = "Number of people") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        panel.grid.minor = element_blank())
```

## Cost analysis

```{r tbl-cost, results='asis'}
cost_tbl <- data.frame(
  `Cost category` = c("Intervention costs", "ART provision", "Total programme cost"),
  Baseline = c(fmt_money(out_base$total_intervention_cost),
               fmt_money(out_base$art_provision_cost),
               paste0("<strong>", fmt_money(out_base$total_cost), "</strong>")),
  `Scenario 1` = c(
    cell(out_s1$total_intervention_cost, diff_s1$diff_intervention_cost, "down", money = TRUE),
    cell(out_s1$art_provision_cost,      diff_s1$diff_art_provision_cost, "down", money = TRUE),
    paste0("<strong>", fmt_money(out_s1$total_cost), "</strong> ",
           fmt_delta(diff_s1$diff_total_cost, "down", money = TRUE))
  ),
  `Scenario 2` = c(
    cell(out_s2$total_intervention_cost, diff_s2$diff_intervention_cost, "down", money = TRUE),
    cell(out_s2$art_provision_cost,      diff_s2$diff_art_provision_cost, "down", money = TRUE),
    paste0("<strong>", fmt_money(out_s2$total_cost), "</strong> ",
           fmt_delta(diff_s2$diff_total_cost, "down", money = TRUE))
  ),
  check.names = FALSE
)
knitr::kable(cost_tbl, format = "html", escape = FALSE,
             align = c("l", "r", "r", "r"),
             table.attr = 'class="report-table"')
```

### Cost by programme area (US$)

```{r cost-breakdown-plot, fig.width=8, fig.height=3.6, out.width="100%"}
library(ggplot2)
library(scales)

cost_cat_levels <- c("ART provision", "Prevention", "Testing",
                     "Treatment monitoring & quality",
                     "Retention & Adherence", "Advanced HIV")
cost_cat_colors <- c(
  "ART provision"                  = "#475569",
  "Prevention"                     = "#10b981",
  "Testing"                        = "#3b82f6",
  "Treatment monitoring & quality" = "#f59e0b",
  "Retention & Adherence"          = "#ec4899",
  "Advanced HIV"                   = "#8b5cf6"
)
pull_cost <- function(o) c(o$art_provision_cost, o$prevention_cost, o$testing_cost,
                           o$treatment_monitoring_cost, o$retention_cost,
                           o$advanced_disease_cost)

cost_bd <- data.frame(
  Scenario = factor(rep(c("Baseline", "Scenario 1", "Scenario 2"),
                        each = length(cost_cat_levels)),
                    levels = c("Baseline", "Scenario 1", "Scenario 2")),
  Category = factor(rep(cost_cat_levels, times = 3), levels = cost_cat_levels),
  Cost     = c(pull_cost(out_base), pull_cost(out_s1), pull_cost(out_s2))
)

cost_bd$bar_total <- as.numeric(tapply(cost_bd$Cost, cost_bd$Scenario, sum)[as.character(cost_bd$Scenario)])
cost_bd$frac      <- ifelse(cost_bd$bar_total > 0, cost_bd$Cost / cost_bd$bar_total, 0)
cost_tot <- data.frame(
  Scenario  = factor(levels(cost_bd$Scenario), levels = levels(cost_bd$Scenario)),
  bar_total = as.numeric(tapply(cost_bd$Cost, cost_bd$Scenario, sum))
)

ggplot(cost_bd, aes(x = Scenario, y = Cost, fill = Category)) +
  # reverse => first level (ART provision) at the BOTTOM (cascade order)
  geom_col(width = 0.62, position = position_stack(reverse = TRUE)) +
  geom_text(aes(label = ifelse(frac >= 0.03, scales::dollar(Cost, accuracy = 1), "")),
            position = position_stack(vjust = 0.5, reverse = TRUE),
            colour = "white", fontface = "bold", size = 2.7) +
  geom_text(data = cost_tot, inherit.aes = FALSE,
            aes(x = Scenario, y = bar_total, label = scales::dollar(bar_total, accuracy = 1)),
            vjust = -0.4, fontface = "bold", size = 3) +
  scale_fill_manual(values = cost_cat_colors, breaks = cost_cat_levels) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.10))) +
  labs(x = NULL, y = "Cost (US$)", fill = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "right", panel.grid.major.x = element_blank())
```

```{r tbl-cost-breakdown, results='asis'}
vb  <- pull_cost(out_base)
v1  <- pull_cost(out_s1)
v2  <- pull_cost(out_s2)
bd_tbl <- data.frame(
  `Cost category` = c(cost_cat_levels, "Total programme cost"),
  Baseline        = c(fmt_money(vb), paste0("<strong>", fmt_money(sum(vb)), "</strong>")),
  `Scenario 1`    = c(fmt_money(v1), paste0("<strong>", fmt_money(sum(v1)), "</strong>")),
  `Scenario 2`    = c(fmt_money(v2), paste0("<strong>", fmt_money(sum(v2)), "</strong>")),
  check.names = FALSE
)
knitr::kable(bd_tbl, format = "html", escape = FALSE,
             align = c("l", "r", "r", "r"),
             table.attr = 'class="report-table"')
```

```{r tbl-prep, results='asis'}
# BY-GROUP entry mode: show where each scenario's PrEP lands across
# FSW/MSM/AGYW/General, with population denominators and coverage.
#
# TOTAL entry mode: the per-group split is produced by split_prep_total(), not
# entered by the user, so the breakdown is suppressed (the entered totals appear
# in "Intervention volumes" instead). The overflow/undelivered notes are the one
# thing NOT suppressed: if part of an entered total could not be delivered, the
# volumes table shows the DELIVERED figure and the gap would otherwise be
# invisible. When nothing overflowed or was discarded there is nothing to say
# and the whole section -- heading and page break included -- is dropped.
prep_note_lines <- function(d) {
  if (is.null(d)) return(character(0))
  out <- character(0)
  if (!is.null(d$overflow_to_general) && d$overflow_to_general > 0.5)
    out <- c(out, sprintf("%s reallocated to the general population (a targeted group reached 100%% coverage)",
                          fmt_n(d$overflow_to_general)))
  if (!is.null(d$discarded) && d$discarded > 0.5)
    out <- c(out, sprintf("%s not delivered &mdash; all groups already at 100%% coverage",
                          fmt_n(d$discarded)))
  out
}

render_prep_block <- function(label, d) {
  if (is.null(d) || is.null(d$table) || nrow(d$table) == 0) return(invisible())
  t <- d$table
  cat(sprintf('<h3 style="font-size:11pt;color:#1F3864;margin:10px 0 2px;">%s</h3>',
              htmltools::htmlEscape(label)))
  cat('<p style="margin:0 0 4px;">Entered directly by group.</p>')
  body <- paste(vapply(seq_len(nrow(t)), function(i) sprintf(
    '<tr><td>%s</td><td>%s</td><td>%s</td><td>%s</td><td>%s</td><td>%s</td></tr>',
    htmltools::htmlEscape(t$Group[i]),
    fmt_n(t[["Oral PrEP"]][i]), fmt_n(t$Lenacapavir[i]), fmt_n(t$Total[i]),
    if (is.na(t$Population[i]))   "&mdash;" else fmt_n(t$Population[i]),
    if (is.na(t[["Coverage %"]][i])) "&mdash;" else sprintf("%.1f%%", t[["Coverage %"]][i])
  ), character(1)), collapse = "\n")
  cat('<table class="report-table">\n',
      '<thead><tr><th>Group</th><th>Oral PrEP</th><th>Lenacapavir</th>',
      '<th>Total</th><th>Population</th><th>Coverage</th></tr></thead>\n<tbody>\n',
      body, '\n</tbody>\n</table>\n', sep = "")
}

scen_labels <- list(baseline = "Baseline", scenario1 = "Scenario 1",
                    scenario2 = "Scenario 2")

if (is.null(pd) || length(pd) == 0) {
  cat('<div class="page-break"></div>\n')
  cat('<h2>PrEP delivery</h2>\n')
  cat("<p><em>PrEP delivery detail unavailable.</em></p>")
} else if (!prep_total) {
  cat('<div class="page-break"></div>\n')
  cat('<h2>PrEP delivery by group</h2>\n')
  for (k in names(scen_labels)) render_prep_block(scen_labels[[k]], pd[[k]])
} else {
  notes <- lapply(names(scen_labels), function(k) prep_note_lines(pd[[k]]))
  names(notes) <- names(scen_labels)
  if (any(lengths(notes) > 0)) {
    cat('<div class="page-break"></div>\n')
    cat('<h2>PrEP allocation notes</h2>\n')
    cat('<p style="margin:0 0 6px;">PrEP was entered as programme totals. ',
        'The volumes shown elsewhere in this report are the amounts ',
        '<strong>delivered</strong>; where an entered total exceeded what the ',
        'eligible population could absorb, the difference is noted below.</p>\n',
        sep = "")
    for (k in names(scen_labels)) {
      if (length(notes[[k]]) == 0) next
      cat(sprintf('<h3 style="font-size:11pt;color:#1F3864;margin:10px 0 2px;">%s</h3>\n',
                  htmltools::htmlEscape(scen_labels[[k]])))
      cat('<ul style="margin:0 0 6px 18px;">',
          paste(sprintf("<li>%s</li>", notes[[k]]), collapse = ""),
          '</ul>\n', sep = "")
    }
  }
}
```

## Intervention volumes by scenario

```{r interv-table, results='asis'}
# Build a long table of every intervention with its baseline/scenario volumes.
rows <- list()
for (gkey in names(ig)) {
  grp <- ig[[gkey]]
  # group header row
  rows[[length(rows) + 1]] <- sprintf(
    '<tr style="background-color:#eef2f7;"><td colspan="4"><strong>%s</strong></td></tr>',
    htmltools::htmlEscape(grp$name)
  )
  # In total-entry mode the eight per-group PrEP keys were never entered by the
  # user -- baseline_input_values() derives them via split_prep_total(). Print
  # two delivered totals instead, positioned where the first PrEP row would
  # have been. Summing the DELIVERED per-group values (rather than echoing the
  # entered total) keeps every number in this table one the model actually used;
  # any shortfall is reconciled in "PrEP allocation notes".
  if (prep_total && any(PREP_GROUP_KEYS %in% names(grp$interventions))) {
    sum_keys <- function(vals, keys)
      sum(vapply(keys, function(k) { v <- vals[[k]]; if (is.null(v)) 0 else as.numeric(v) },
                 numeric(1)), na.rm = TRUE)
    for (tk in list(list(nm = "Oral PrEP (programme total, all groups)", keys = PREP_ORAL_KEYS),
                    list(nm = "Lenacapavir (programme total, all groups)", keys = PREP_LENA_KEYS))) {
      rows[[length(rows) + 1]] <- sprintf(
        "<tr><td>%s</td><td>%s</td><td>%s</td><td>%s</td></tr>",
        htmltools::htmlEscape(tk$nm),
        fmt_n(sum_keys(base_vals, tk$keys)),
        fmt_n(sum_keys(s1_vals,   tk$keys)),
        fmt_n(sum_keys(s2_vals,   tk$keys))
      )
    }
  }
  for (ikey in names(grp$interventions)) {
    if (prep_total && ikey %in% PREP_GROUP_KEYS) next
    int <- grp$interventions[[ikey]]
    bv  <- base_vals[[ikey]]; if (is.null(bv)) bv <- 0
    s1v <- s1_vals[[ikey]];   if (is.null(s1v)) s1v <- 0
    s2v <- s2_vals[[ikey]];   if (is.null(s2v)) s2v <- 0
    # clinical_visit_12month is typed "coverage" so it travels on the same
    # numeric 0/100 contract as the other coverage inputs, but it is a binary
    # 6-vs-12-month choice (radioGroupButtons in the app, choiceValues 0/100).
    # Printing it as "100.0%" is meaningless to a report reader, so render the
    # choice label instead. >= 50 is the same threshold the app uses to pick
    # the selected button.
    is_pct <- isTRUE(int$type == "coverage")
    fmt_v <- if (identical(ikey, "clinical_visit_12month")) {
      function(x) {
        xn <- suppressWarnings(as.numeric(x))
        if (!isTRUE(is.finite(xn))) "&mdash;"
        else if (xn >= 50) "Every 12 months" else "Every 6 months"
      }
    } else if (is_pct) {
      function(x) sprintf("%.1f%%", x)
    } else {
      fmt_n
    }
    rows[[length(rows) + 1]] <- sprintf(
      "<tr><td>%s</td><td>%s</td><td>%s</td><td>%s</td></tr>",
      htmltools::htmlEscape(int$name),
      fmt_v(bv), fmt_v(s1v), fmt_v(s2v)
    )
  }
}
cat('<table>\n<thead>',
    '<tr><th>Intervention</th><th>Baseline</th><th>Scenario 1</th><th>Scenario 2</th></tr>',
    '</thead>\n<tbody>\n',
    paste(rows, collapse = "\n"),
    '\n</tbody>\n</table>',
    sep = "")
```

```{r appendix-params, results='asis'}
# APPENDIX -- model parameters changed from defaults.
#
# Omitted ENTIRELY when nothing was changed (Alex's call): a run at defaults
# says so by the absence of the appendix, not by an empty table. param_overrides
# arrives as a zero-row data.frame in that case, built by
# param_override_summary() in the app file -- which is the single arbiter of
# what counts as "changed". Do not re-derive that here.
#
# Grouped by Section with a spanning header row, matching the intervention
# volumes table above. Values arrive pre-formatted in the units the Parameters
# tab displays (percent / months / USD), so this chunk does no conversion.
has_overrides <- !is.null(po) && is.data.frame(po) && nrow(po) > 0

if (has_overrides) {
  cat('<div class="page-break"></div>\n')
  cat('<h2>Appendix &mdash; model parameters changed from defaults</h2>\n')
  cat('<p style="margin:0 0 8px;">The results in this report were produced with ',
      'the parameter values below in place of the tool&rsquo;s defaults. ',
      'All parameters not listed here were left at their default values.</p>\n',
      sep = "")
  arows <- character(0)
  for (sec in unique(po$Section)) {
    arows <- c(arows, sprintf(
      '<tr style="background-color:#eef2f7;"><td colspan="3"><strong>%s</strong></td></tr>',
      htmltools::htmlEscape(sec)))
    sub <- po[po$Section == sec, , drop = FALSE]
    for (i in seq_len(nrow(sub))) {
      arows <- c(arows, sprintf(
        "<tr><td>%s</td><td>%s</td><td>%s</td></tr>",
        htmltools::htmlEscape(sub$Parameter[i]),
        htmltools::htmlEscape(sub$Default[i]),
        sprintf("<strong>%s</strong>", htmltools::htmlEscape(sub[["Set to"]][i]))))
    }
  }
  cat('<table>\n<thead>',
      '<tr><th>Parameter</th><th>Default</th><th>Used in this report</th></tr>',
      '</thead>\n<tbody>\n', paste(arows, collapse = "\n"),
      '\n</tbody>\n</table>\n', sep = "")
}
```

<div class="footer-note">
Use the tool to compare directions and trade-offs. Use country processes for the exact numbers in your plan. Report generated by the
TIER-Plus Tool.
</div>
